---
title: "三元图绘制"
author: "刘永鑫(Yong-Xin Liu)"
date: "`r Sys.Date()`"
output: html_document
---
 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
 
## 设置输入输出文件和参数

绘制三元图，是三组比较；本质上是三组间两两比较再比较。

```{r}

#############################################################
# Title: Didfferent abundance OTU -- Tenary plot
# Author: Yong-Xin Liu
# E-mail: liuyongxin@caas.cn
# Date: 2017/6/21
# Update: 2021/4/3
# Description: Script to draw tenary plot
# Input File required in the blow list: 
# 需要DA_OTU计算各组间的差异OTU结果
# 1. "metadata.txt"  Design of experiment
# 2. "otu_table.txt" OTU table, better include taxonomy
# 3. "tern.txt" ternary list
# Group A/B/C lie in left, right and top of triangle
#############################################################

# Set working enviroment
rm(list=ls()) # clean enviroment object
source("ternary_func.R")

if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
if (!require("Biobase", quietly = TRUE))
    BiocManager::install("Biobase")

# Rewrite header of OTU count
write.table(paste("GroupA-Left","GroupB-Right","GroupC-Top/Common-Middle",sep = "\t"), file=paste("ter_", "sum.txt", sep=""), sep="\t", quote=F, append = FALSE, row.names=F, col.names=F)

# default color:left green, right orange, top red, background grey; alternative to yellow, blue, dark_green, very_dark_green, sea_green, black, dark_brown, drak_red
color=c(c_green,c_orange,c_red,c_grey) 

# Read the design & OTU table with taxonomy information
design = read.table("metadata.txt", header=T, row.names= 1, sep="\t") 
otu_table = read.delim("otutab.txt", row.names= 1,  header=T, sep="\t")
tern_group=read.table("tern.txt", header=F, sep="\t") 

# loop for each ternery plot
for(i in 1:dim(tern_group)[1]){
  # i=1
  # Format a line to a column, then can format to character
  tern=as.character(t(tern_group[i,]))
  
  # Select design and otu_table
  if (TRUE){sub_design = subset(design,group %in% tern )}else{sub_design=design}

  # 比较实验设计与OTU表一致；因为样品会有缺失
  idx = rownames(sub_design) %in% colnames(otu_table) 
  sub_design = sub_design[idx,]
  count = otu_table[, rownames(sub_design)] 

  # normalize to 1000
  norm= as.data.frame(t(t(count)/colSums(count))*100)
  colSums(norm)
  
  # merge group to mean
  ## 按样品名合并实验组与转置的OTU
  mat_t2 = merge(sub_design[c("group")], t(norm), by="row.names")[,-1]
  ## 按实验设计求组平均值
  mat_mean = aggregate(mat_t2[,-1], by=mat_t2[1], FUN=mean) # mean 
#  per3= do.call(rbind, mat_mean)[-1,] # Error in rowSums(per3) : 'x' must be numeric
  per3=t(mat_mean[,-1])
  colnames(per3) = mat_mean$group
  per3=as.data.frame(per3[rowSums(per3)>0,]) # remove all 0 OTU
  per3=per3[,tern] # reorder per3 as input

  # 三元图中A/B/C三个组分别位于三角形的左、右和顶
  # 1. 三角相对于另两者显著富集的OTU
  per3$color=color[4] # set all default # 设置点默认颜色为灰
  ## select A specific
  AvB = read.table(paste("otu_",tern[1],"vs",tern[2],"_enriched.txt",sep=""), header=T, row.names= 1, sep="\t")
  AvC = read.table(paste("otu_",tern[1],"vs",tern[3],"_enriched.txt",sep=""), header=T, row.names= 1, sep="\t")
  A = intersect(row.names(AvB), row.names(AvC))
  if (length(A)>0){per3[A,]$color=color[1]} # set all default
  ## select B specific
  BvA = read.table(paste("otu_",tern[2],"vs",tern[1],"_enriched.txt",sep=""), header=T, row.names= 1, sep="\t")
  BvC = read.table(paste("otu_",tern[2],"vs",tern[3],"_enriched.txt",sep=""), header=T, row.names= 1, sep="\t")
  B = intersect(row.names(BvA), row.names(BvC))
  if (length(B)>0){per3[B,]$color=color[2]} # set all default
  ## select C specific
  CvA = read.table(paste("otu_",tern[3],"vs",tern[1],"_enriched.txt",sep=""), header=T, row.names= 1, sep="\t")
  CvB = read.table(paste("otu_",tern[3],"vs",tern[2],"_enriched.txt",sep=""), header=T, row.names= 1, sep="\t")
  C = intersect(row.names(CvA), row.names(CvB))
  if (length(C)>0){per3[C,]$color=color[3]} # set all default
  ## output pdf and png in 8x8 inches
  per3lg=log2(per3[,1:3]*1+1) # 对数变换，差OTU千分比的差距，点大小更均匀
  pdf(file=paste("ter_",tern[1],tern[2],tern[3],".pdf", sep=""), height = 8, width = 8)
  tern_e(per3lg[,1:3], prop=T, col=per3$color, grid_color="black", labels_color="transparent", pch=19, main="Tenary Plot")
  dev.off()
  # png(file=paste("ter_",tern[1],tern[2],tern[3],".png", sep=""), height = 8, width = 8, units = "in", res = 300)
  # tern_e(per3lg[,1:3], prop=T, col=per3$color, grid_color="black", labels_color="transparent", pch=19, main="Tenary Plot")
  # dev.off()
  ## Write hightlight OUT
  if (length(A)>0){write.table(AvB[A,], file=paste("ter_", tern[1],tern[2],tern[3],tern[1],".txt", sep=""), sep="\t", quote=F, row.names=T, col.names=T)}
  if (length(B)>0){write.table(BvA[B,], file=paste("ter_", tern[1],tern[2],tern[3],tern[2],".txt", sep=""), sep="\t", quote=F, row.names=T, col.names=T)}
  if (length(C)>0){write.table(CvA[C,], file=paste("ter_", tern[1],tern[2],tern[3],tern[3],".txt", sep=""), sep="\t", quote=F, row.names=T, col.names=T)}
  ## Write hightlight OUT summary
  write.table(paste(tern[1],tern[2],tern[3],sep = "\t"), file=paste("ter", "_sum.txt", sep=""), sep="\t", quote=F, append = TRUE, row.names=F, col.names=F)
  write.table(paste(length(A),length(B),length(C),sep = "\t"), file=paste("ter", "_sum.txt", sep=""), sep="\t", quote=F, append = TRUE, row.names=F, col.names=F)
  
  # 2. 两底角相对于顶点显著富集的OTU，分共有和特有，类似维恩图
  per3$color=color[4] # set all default # 设置点默认颜色为灰
  AvC = read.table(paste("otu_",tern[1],"vs",tern[3],"_enriched.txt",sep=""), header=T, row.names= 1, sep="\t")
  BvC = read.table(paste("otu_",tern[2],"vs",tern[3],"_enriched.txt",sep=""), header=T, row.names= 1, sep="\t")
  C = intersect(row.names(AvC), row.names(BvC))
  A = setdiff(row.names(AvC), C) 
  B = setdiff(row.names(BvC), C) 
  if (length(A)>0){per3[A,]$color=color[1]} 
  if (length(B)>0){per3[B,]$color=color[2]} 
  if (length(C)>0){per3[C,]$color=color[3]}
  ## output pdf and png in 8x8 inches
  per3lg=log2(per3[,1:3]*1+1) # 对数变换，差OTU千分比的差距，点大小更均匀
  pdf(file=paste("ter_",tern[1],tern[2],tern[3],"venn.pdf", sep=""), height = 8, width = 8)
  tern_e(per3lg[,1:3], prop=T, col=per3$color, grid_color="black", labels_color="transparent", pch=19, main="Tenary Plot")
  dev.off()
  # png(file=paste("ter_",tern[1],tern[2],tern[3],"venn.png", sep=""), height = 8, width = 8, units = "in", res = 300)
  # tern_e(per3lg[,1:3], prop=T, col=per3$color, grid_color="black", labels_color="transparent", pch=19, main="Tenary Plot")
  # dev.off()
  ## Write hightlight OUT
  if (length(A)>0){write.table(AvC[A,], file=paste("ter_", tern[1],tern[2],tern[3],tern[1],"venn.txt", sep=""), sep="\t", quote=F, row.names=T, col.names=T)}
  if (length(B)>0){write.table(BvC[B,], file=paste("ter_", tern[1],tern[2],tern[3],tern[2],"venn.txt", sep=""), sep="\t", quote=F, row.names=T, col.names=T)}
  if (length(C)>0){write.table(AvC[C,], file=paste("ter_", tern[1],tern[2],tern[3],tern[3],"venn.txt", sep=""), sep="\t", quote=F, row.names=T, col.names=T)}
  ## Write hightlight OUT summary
  write.table(paste(tern[1],tern[2],"Common",sep = "\t"), file=paste("ter", "_sum.txt", sep=""), sep="\t", quote=F, append = TRUE, row.names=F, col.names=F)
  write.table(paste(length(A),length(B),length(C),sep = "\t"), file=paste("ter", "_sum.txt", sep=""), sep="\t", quote=F, append = TRUE, row.names=F, col.names=F)
  
}

```


使用此脚本，请引用下文：

If used this script, please cited:

**Yong-Xin Liu**, Lei Chen, Tengfei Ma, Xiaofang Li, Maosheng Zheng, Xin Zhou, Liang Chen, Xubo Qian, Jiao Xi, Hongye Lu, Huiluo Cao, Xiaoya Ma, Bian Bian, Pengfan Zhang, Jiqiu Wu, Ren-You Gan, Baolei Jia, Linyang Sun, Zhicheng Ju, Yunyun Gao, **Tao Wen**, **Tong Chen**. 2023. EasyAmplicon: An easy-to-use, open-source, reproducible, and community-based pipeline for amplicon data analysis in microbiome research. **iMeta** 2: e83. https://doi.org/10.1002/imt2.83

Copyright 2016-2023 Yong-Xin Liu <liuyongxin@caas.cn>, Tao Wen <taowen@njau.edu.cn>, Tong Chen <chent@nrc.ac.cn>
